---
title: "fit_Zscore_to_T"
output: ht  vml_document
date: "2023-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("MASS")
library(MASS)
#install.packages("fGarch")
library(fGarch)
library(dplyr)
library(moments) 
library(sn) 
library(ggplot2)
library(gridExtra)
```
## R Markdown

```{r}
df1 <- read.csv("/data2/stan/quickBEAST/a8.789625_b8.789625/lambda0.04545/parametrized/ASE_0.05_error/g-1000_h-10_d-100_t-1.txt",sep="\t")
X1 <- df1$qb_zscore
# Check for NA values
sum(is.na(X1))
```


```{r}
hist(X1)
```

```{r}
fit_t_distribution <- function(data) {
  fit <- fitdistr(data, "t")
  return(fit)
}
fit_skewed_t_distribution <- function(data) {
  fit <- sstdFit(data)
  return(fit)
}
fit_normal_distribution <- function(data) {
  mean_value <- mean(data, na.rm = TRUE)
  sd_value <- sd(data, na.rm = TRUE)
  fit <- list(mean = mean_value, sd = sd_value)
  return(fit)
}

# Function to generate samples from a t-distribution
generate_t_samples <- function(fit, n) {
  rt(n, df = fit$estimate["df"])
}

generate_skewed_t_samples <- function(fit, n) {
  # Extract parameters from the fit object
  mean <- fit$estimate["mean"]
  sd <- fit$estimate["sd"]
  xi <- fit$estimate["xi"]
  nu <- fit$estimate["nu"]

  # Check for valid parameters
  if(is.na(mean) || is.na(sd) || is.na(xi) || is.na(nu) || nu <= 0) {
    stop("Invalid or NA values in fitted parameters.")
  }

  # Generate samples
  rsstd(n, mean = mean, sd = sd, xi = xi, nu = nu)
}

generate_normal_samples <- function(fit, n) {
  rnorm(n, mean = fit$mean, sd = fit$sd)
}


```

```{r}
# Fit normal, t, and skewed t distributions for X1
fit_normal_X1 <- fit_normal_distribution(X1)
fit_t_X1 <- fit_t_distribution(X1)
fit_skewed_t_X1 <- fit_skewed_t_distribution(X1)

# Generate samples from each fitted distribution
normal_samples_X1 <- generate_normal_samples(fit_normal_X1, length(X1))
t_samples_X1 <- generate_t_samples(fit_t_X1, length(X1))
skewed_t_samples_X1 <- generate_skewed_t_samples(fit_skewed_t_X1, length(X1))
```

```{r}
hist(X1, breaks = 100)
```
```{r}
hist(normal_samples_X1, breaks = 100)
```

```{r}
hist(t_samples_X1, breaks = 100)
```

```{r}
hist(skewed_t_samples_X1, breaks = 100)
```

```{r}
# Combine all data and labels
data_list <- list(
  data.frame(Value = X1, Group = "Original"),
  data.frame(Value = normal_samples_X1, Group = "Normal"),
  data.frame(Value = t_samples_X1, Group = "T-Distribution"),
  data.frame(Value = skewed_t_samples_X1, Group = "Skewed T-Distribution")
)

labels <- c("Original", "Normal", "T-Distribution", "Skewed T-Distribution")

# Create a function to plot overlaid histograms
plot_overlaid_histograms <- function(data, labels) {
  ggplot(data, aes(x = Value, fill = Group)) +
    geom_histogram(alpha = 0.5, position = "identity", bins = 30) +
    labs(
      title = "Overlaid Histograms",
      x = "Value",
      y = "Frequency",
      fill = "Legend Title"
    ) +
    scale_fill_manual(values = c("Original" = "black", "Normal" = "red", "T-Distribution" = "green", "Skewed T-Distribution" = "blue")) +
    theme(
      panel.background = element_rect(fill = "transparent"),
      text = element_text(size = 14),  # Adjust font size for text
      legend.text = element_text(size = 12),  # Adjust font size for legend text
      legend.title = element_text(size = 14)  # Adjust font size for legend title
    )
}

# Create a function to plot overlaid density plots
plot_overlaid_density <- function(data, labels) {
  ggplot(data, aes(x = Value, color = Group)) +
    geom_density(alpha = 0.5) +
    labs(
      title = "Overlaid Density Plots",
      x = "Value",
      y = "Density",
      color = "Legend Title"
    ) +
    scale_color_manual(values = c("Original" = "black", "Normal" = "red", "T-Distribution" = "green", "Skewed T-Distribution" = "blue")) +
    theme(
      panel.background = element_rect(fill = "transparent"),
      text = element_text(size = 14),  # Adjust font size for text
      legend.text = element_text(size = 12),  # Adjust font size for legend text
      legend.title = element_text(size = 14)  # Adjust font size for legend title
    )
}

# Combine both plots side by side
combined_plots <- grid.arrange(
  plot_overlaid_histograms(do.call(rbind, data_list), labels),
  plot_overlaid_density(do.call(rbind, data_list), labels),
  ncol = 2
)

# Print the combined plots
print(combined_plots)

```

```
```{r}
plot_histograms <- function(original_data, t_samples, skewed_t_samples) {
  # Set the range for the histogram
  combined_data <- c(original_data, t_samples, skewed_t_samples)
  data_range <- range(combined_data)

  # Plot histogram of the original data
  hist(original_data, freq = FALSE, xlim = data_range, 
       main = "Histogram Comparison", xlab = "Value", col = rgb(0, 0, 1, 0.25), breaks = 30)

  # Add histogram of t-distribution samples
  hist(t_samples, freq = FALSE, add = TRUE, col = rgb(0, 1, 0, 0.25), breaks = 30)

  # Add histogram of skewed t-distribution samples
  hist(skewed_t_samples, freq = FALSE, add = TRUE, col = rgb(1, 1, 0, 0.25), breaks = 30)

  # Add a legend
  legend("topright", legend = c("Original", "T-Distribution", "Skewed T-Distribution"),
         fill = c(rgb(0, 0, 1, 0.25), rgb(0, 1, 0, 0.25), rgb(1, 1, 0, 0.25)))
}
# You can then plot histograms or conduct further analysis as needed
plot_histograms(X1, t_samples_X1, skewed_t_samples_X1)
```


```{r fig.width=8, fig.height=6}
plot_distributions <- function(original_data, normal_samples, t_samples, skewed_t_samples) {
  plot_density <- function(data, col, add=FALSE) {
    density_data <- density(data)
    if (!add) {
      plot(density_data, main="Distribution Comparison", xlab="Value", ylab="Density", 
           col=col, xlim=range(density_data$x), ylim=range(0, density_data$y), lwd=2)
    } else {
      lines(density_data, col=col, lwd=2)
    }
  }

  # Plot original data density
  plot_density(original_data, "black")

  # Add normal distribution density
  plot_density(normal_samples, "red", add=TRUE)

  # Add t-distribution density
  plot_density(t_samples, "blue", add=TRUE)

  # Add skewed t-distribution density
  plot_density(skewed_t_samples, "green", add=TRUE)

  # Add a legend
  legend("topright", legend = c("Original", "Normal", "T-Distribution", "Skewed T-Distribution"),
         col = c("black", "red", "blue", "green"), lwd=2)
}

# Usage:
plot_distributions(X1, normal_samples_X1, t_samples_X1, skewed_t_samples_X1)

```

```{r fig.width=8, fig.height=6}
plot_distributions <- function(original_data, t_samples, skewed_t_samples) {
  plot_density <- function(data, col, add=FALSE) {
    density_data <- density(data)
    if (!add) {
      plot(density_data, main="Distribution Comparison", xlab="Value", ylab="Density", 
           col=col, xlim=range(density_data$x), ylim=range(0, density_data$y), lwd=2)
    } else {
      lines(density_data, col=col, lwd=2)
    }
  }

  # Plot original data density
  plot_density(original_data, "black")


  # Add t-distribution density
  plot_density(t_samples, "blue", add=TRUE)

  # Add skewed t-distribution density
  plot_density(skewed_t_samples, "green", add=TRUE)

  # Add a legend
  legend("topright", legend = c("Original", "T-Distribution", "Skewed T-Distribution"),
         col = c("black", "blue", "green"), lwd=2)
}

# Usage:
plot_distributions(X1, t_samples_X1, skewed_t_samples_X1)

```
#### Z-scores against n/t/st distribution
```{r fig.width=6, fig.height=6}
# Load necessary libraries
library(sn)
library(ggplot2)

# Set the random seed for reproducibility
set.seed(123)

data <- X1
# Sort the data
sorted_data <- sort(data)
```

```{r fig.width=8, fig.height=6}
plot_qq <- function(qq_plot_data,theoretical_quantiles,title){
  # Calculate the negative log10 of the quantiles
  qq_plot_data$log10_Observed <- -log10(qq_plot_data$Observed)
  qq_plot_data$log10_Theoretical <- -log10(theoretical_quantiles)
  # Create a QQ plot
  # Create a QQ plot with the same x-axis and y-axis limits
  ggplot(qq_plot_data, aes(x = log10_Theoretical, y = log10_Observed)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
    labs(title = paste0("QQ Plot in -log10 Scale against ",title," Distribution",sep=""),
         x = "-log10(Theoretical Quantiles)",
         y = "-log10(Observed Quantiles)") +
    coord_fixed(ratio = 1)+
        xlim(c(-1,4)) +   # Set the same x-axis limits
    ylim(c(-1,4)) +
    theme_minimal() +
    theme(axis.text = element_text(size = 12),  # Increase font size for axis labels
          axis.title = element_text(size = 14), # Increase font size for axis titles
          plot.title = element_text(size = 16)) # Increase font size for t
}

plot_qq <- function(qq_plot_data,theoretical_quantiles,title){
  # Create a QQ plot
  # Create a QQ plot with the same x-axis and y-axis limits
  ggplot(qq_plot_data, aes(x = theoretical_quantiles, y = qq_plot_data$Observed)) +
    geom_point() +
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
    labs(title = paste0("QQ Plot against ",title," Distribution",sep=""),
         x = "Theoretical Quantiles",
         y = "Observed Quantiles") +
    coord_fixed(ratio = 1)+
        xlim(c(-1,4)) +   # Set the same x-axis limits
    ylim(c(-1,4)) +
    theme_minimal() +
    theme(axis.text = element_text(size = 12),  # Increase font size for axis labels
          axis.title = element_text(size = 14), # Increase font size for axis titles
          plot.title = element_text(size = 16)) # Increase font size for t
}
qq_normal <- function(data,title){
  sorted_data<-sort(data)
  # Create a QQ plot in -log10 scale against a normal distribution
  qq_plot_data <- data.frame(Observed = sorted_data)
  theoretical_quantiles <- qnorm(ppoints(length(sorted_data)))  # Calculate theoretical quantiles
  plot_qq(qq_plot_data,theoretical_quantiles,title)
}
qq_t <- function(data,title){
  sorted_data<-sort(data)
  # Create a QQ plot in -log10 scale against a normal distribution
  qq_plot_data <- data.frame(Observed = sorted_data)
  theoretical_quantiles <- qt(ppoints(length(sorted_data)),df=df)  # Adjust df as needed
  plot_qq(qq_plot_data,theoretical_quantiles,title)
}


qq_st <- function(data, title) {
  # Fit your data to a skewed t-distribution using sstdFit
  fit_result <- sstdFit(data)
  
  # Extract estimated parameters, including skewness
  skewness_est <- fit_result$gradient["xi"]

  sorted_data <- sort(data)
  n <- length(sorted_data)
  
  # Calculate quantiles from a theoretical skewed t-distribution
  theoretical_quantiles <- qsn(ppoints(n), xi = skewness_est)
  
  # Create a QQ plot in -log10 scale against a normal distribution
  qq_plot_data <- data.frame(Observed = sorted_data)
  plot_qq(qq_plot_data, theoretical_quantiles, title)
}
```


```{r}
qq_normal(X1,"normal")
```


```{r}
qq_t(X1,"t")
```

```{r}
qq_st(X1,"skewed t")
```

